---
title: "Configuration File"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Configuration File}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org"))
load(here::here("data", "configuration.rda"))
```

# NWFSC West Coast Groundfish Bottom Trawl Survey

The below configurations were used in the most recent assessment:

```{r, echo=FALSE}
configuration |>
  dplyr::filter(source == "NWFSC.Combo", used == TRUE) |>
  dplyr::select(-fxn, -source, -formula) |>
  gt::gt() |>
  gt::tab_options(
    table.font.size = 10
  )
```

# NWFSC Slope Survey

The below configurations were used in the most recent assessment:

```{r, echo=FALSE}
configuration |>
  dplyr::filter(source == "NWFSC.Slope", used == TRUE) |>
  dplyr::select(-fxn, -source, -formula) |>
  gt::gt() |>
  gt::tab_options(
    table.font.size = 10
  )
```

# AFSC Slope Survey

The below configurations were used in the most recent assessment:

```{r, echo=FALSE}
configuration |>
  dplyr::filter(source == "AFSC.Slope", used == TRUE) |>
  dplyr::select(-fxn, -source, -formula) |>
  gt::gt() |>
  gt::tab_options(
    table.font.size = 10
  )
```


# AFSC/NWFSC Triennial Shelf Survey

The below configurations were used in the most recent assessment:

```{r, echo=FALSE}
configuration |>
  dplyr::filter(source == "Triennial", used == TRUE) |>
  dplyr::select(-fxn, -source, -formula) |>
  gt::gt() |>
  gt::tab_options(
    table.font.size = 10
  )
```

# Other Configurations

The below configurations are alternative configurations that were not used in the most recent assessment:

```{r, echo=FALSE}
configuration |>
  dplyr::filter(used == FALSE) |>
  dplyr::select(-fxn, -formula) |>
  gt::gt() |>
  gt::tab_options(
    table.font.size = 10
  )
```

