---
title: "Coastwide indices"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Coastwide indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```

## Overview

This vignette highlights the ability to generate coastwide indices for stocks using the `indexwc` package.

## Yellowtail rockfish example

As a case study, we'll focus on yellowtail rockfish

```{r}
library(dplyr)
library(indexwc)
library(purrr)
library(sdmTMB)
library(tibble)
```

First we'll load in the configuration file. This is bringing in the data from Github -- but for users who download the source code, this is also available in the `data-raw` folder. 
```{r}
url <- "https://raw.githubusercontent.com/pfmc-assessments/indexwc/refs/heads/main/data-raw/configuration.csv"
configuration <- tibble::as_tibble(read.csv(url))

configuration <- configuration |>
  dplyr::filter(species == "yellowtail rockfish")
```

The configuration file contains settings you may wish to change, including attributes of filters (years, latitudes, depths), and model components (anisotropy, knots, spatiotemporal effects, etc).


### Statistical model

For most indices developed on the west coast of the USA, the main formula is usually something similar to the following, where `fyear` represents a fixed year effect, and `pass_scaled` is factor describing whether the haul is part of pass 1 (early season) or pass 2 (late season).

```{r}
formula <- "catch_weight ~ 0 + fyear + pass_scaled"
```

### Preparing the data

This block pulls trawl survey from the West Coast Groundfish Bottom Trawl Survey and applies filters (based on latitude, depth, and year) to each observation.

```{r}
data <- configuration |>
  dplyr::rowwise() |>
  dplyr::mutate(
    data_raw = list(format_data(eval(parse(text = fxn)))),
    data_filtered = list(data_raw |>
      dplyr::filter(
        depth <= min_depth, depth >= max_depth,
        latitude >= min_latitude, latitude <= max_latitude,
        year >= min_year, year <= max_year
      ))
  ) |>
  dplyr::ungroup()
```

### Fitting the model with indexwc and sdmTMB

Now we can use the `indexwc` package to fit the model. The `indexwc` package acts as a wrapper for `sdmTMB` here, combining the estimation process (`sdmTMB()`) with the index generation (`get_index()`).

This code is optimized to fit multiple models (across stocks / species, or model configurations). In our case, the configurations dataframe has 1 row -- but this could include multiple rows. A key point to highlight is that in this code, we use `sdmTMBcontrol()` to pass in a list that contains the variables above we use to map off and initialize the parameters. For datasets without missing values, you may not want to include this `sdmTMBcontrol()` line, but it may be useful to adjust the `newton_loops`, etc.

Note: this code will fit a single model for each row 

```{r}
best <- data |>
  dplyr::mutate(
    # Evaluate the call in family
    family = purrr::map(family, .f = ~ eval(parse(text = .x))),
    # Run the model on each row in data
    results = purrr::pmap(
      .l = list(
        data = data_filtered,
        formula = formula,
        family = family,
        anisotropy = anisotropy,
        n_knots = knots,
        share_range = share_range,
        spatiotemporal = purrr::map2(spatiotemporal1, spatiotemporal2, list)
      ),
      .f = indexwc::run_sdmtmb
    )
  )
```

### Examining the index and diagnostics

Output and diagnostics for all models and predictions are stored in a folder that `indexwc` creates; it's location is "yellowtail_rockfish/wcgbts/delta_gamma..". The "data" folder contains the raw data used for estimation (`data.rdata`) and a plot of the SPDE mesh is in `mesh.png`. The "index" folder contains diagnostics and predictions for 10 sets of indices. These 10 indices represent (1) N/S of Cape Mendocino, (2) N/S of Monterrey, (3) N/S of Pt Conception, a coastwide index, and state-specific indices. Plots include:
* density plots (`density01.png`, `density02.png`, ...) which represent predictions for pairs of years 
* a coastwide index (`index_coastwide.png`)
* fixed effect parameters (`parameters_fixed_effects.png`)
* QQ plot (`qq.png`)
* residual plots (`residuals_1_01.png`, `residuals_1_02.png`, ..., `residuals_2_01.png`, `residuals_2_02.png`,...) where pairs of years are shown, and separate residual plots are made for the presence-absence and positive model components

Additional output includes
* the dataframe of `sdmTMB::sanity()` checks (in `sanity_data_frame.csv`) 
* indices generated by area (`est_by_area.csv`)
