---
title: "Coastwide indices: simple example"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Coastwide indices: simple example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
options(repos = c(CRAN = "https://cloud.r-project.org"))
```

## Overview

This vignette highlights the ability to generate coastwide indices for stocks using available functions in the `indexwc` package for a West Coast groundfish stock for the West Coast Groundfish Bottom Trawl Survey using the configuration data object.

```{r packages, eval=TRUE, echo=TRUE}
library(dplyr)
library(indexwc)
options(repos = c(CRAN = "https://cloud.r-project.org"))
```

## Get data and settings

```{r config, eval=TRUE, echo=TRUE}
# Load from the rda file
configuration_to_run <- configuration |>
  dplyr::filter(species == "sablefish", source == "NWFSC.Combo", family == "sdmTMB::delta_gamma()") 
```

```{r pull-data, eval = TRUE, echo = TRUE}
my_data <- pull_and_format_data(
   configuration_to_run = configuration_to_run
)
```

## Estimate the model

```{r settings, eval = TRUE, echo = TRUE}
names(my_data)
my_data$family
my_data$formula
my_data$knots
my_data$share_range
my_data$anisotropy
list(my_data$spatiotemporal1, my_data$spatiotemporal2)

```

```{r run-model, eval=FALSE, echo = TRUE}
fit <- run_sdmtmb(
  data = my_data$data_filtered[[1]],
  family = my_data$family,
  formula = my_data$formula,
  n_knots = my_data$knots,
  share_range = my_data$share_range,
  anisotropy = my_data$anisotropy,
  spatiotemporal = list(my_data$spatiotemporal1, my_data$spatiotemporal2)
)
```

## Diagnostics

```{r diagnostics, eval = FALSE, echo = TRUE}
diag <- diagnose(
 fit = fit
)
```

## Calculate an index of abundance

```{r index, eval = FALSE, echo = TRUE}
index <- calc_index_areas(
   data = fit$data,
   fit = fit,
   boundaries = "Coastwide"
)
```

## Write out all files and figures

```{r save-output, eval = FALSE, echo = TRUE}
save_index_outputs(
 fit = fit,
 diagnostics = diag,
 indices = index,
 dir = "C:/indices"
)
```
