---
title: "Coastwide indices: simple example"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Coastwide indices: simple example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
options(repos = c(CRAN = "https://cloud.r-project.org"))
```

## Overview

This vignette highlights the ability to generate coastwide indices for West Coast groundfish stocks for bottom trawl surveys using available functions in the `indexwc` package and the available configuration data object.   


```{r packages, eval=TRUE, echo=TRUE}
library(dplyr)
library(indexwc)
options(repos = c(CRAN = "https://cloud.r-project.org"))
```

## Get data and settings

`indexwc` includes a maintained configuration file that contains default settings for index creation for West Coast groundfish species. The configuration is a data frame with entries by row for species, alternative model error structures (e.g., delta-lognormal, delta-gamma), formulas, and various bottom trawl surveys (e.g., West Coast Groundfish Bottom Trawl Survey [WCGBTS], etc.). The data frame includes various columns defining the commonly used model configuration parameters (e.g., spatiotemporal settings, anisotropy, etc.). Please see the configuration vignette for additional information. 

In the below example, the configuration file will be used to pull and run a model for sablefish using data from the WCGBTS (aka "NWFSC.Combo" as defined in the `nwfscSurvey`) using a delta-gamma model error structure.  

```{r config, eval = TRUE, echo = TRUE}
# Load from the rda file
configuration_to_run <- configuration |>
  dplyr::filter(species == "sablefish", source == "NWFSC.Combo", family == "sdmTMB::delta_gamma()") 

print(configuration_to_run)
```

The selected configuration is then passed to `pull_and_format_data()` that pulls and returns a data list with entries for model settings based on the configuration file. 

```{r pull-data, eval = TRUE, echo = TRUE}
my_data <- pull_and_format_data(
   configuration_to_run = configuration_to_run
)
```

Double check the items in the `my_data` list created above:

```{r settings, eval = TRUE, echo = TRUE}
names(my_data)
my_data$family
my_data$formula
my_data$knots
my_data$share_range
my_data$anisotropy
list(my_data$spatiotemporal1, my_data$spatiotemporal2)

```

## Estimate the model

Running `sdmTMB` models can be done through `run_sdmtmb()` that has function arguments defining the model type and structure to run.  `run_sdmtmb()` includes an argument for saving the fit object from `sdmTMB`, however, the default setting is `dir = NULL` where output will not be saved at this stage.  

```{r run-model, eval = FALSE, echo = TRUE}
fit <- run_sdmtmb(
  data = my_data$data_filtered[[1]],
  family = my_data$family,
  formula = my_data$formula,
  n_knots = my_data$knots,
  share_range = my_data$share_range,
  anisotropy = my_data$anisotropy,
  spatiotemporal = list(my_data$spatiotemporal1, my_data$spatiotemporal2)
)
```

`run_sdmtb()` returns a list that contains all information from running the model.  To see the estimated parameters for the model, type the assigned R object from `run_sdmtmb()`. 

```{r print-fit, eval = FALSE, echo = TRUE}
fit
```


## Diagnostics

A range of diagnostics are available and all diagnostics can be done using `diagnose()`:

```{r diagnostics, eval = FALSE, echo = TRUE}
diag <- diagnose(
 fit = fit
)
```

A sanity check is performed to evaluate model convergence and the AIC is calculated and the fixed effects are evaluated. A prediction grid based upon the California Current is created within `diagnose()` that is then used to evaluate QQ, residuals, and model predicted density.  

`diagnose()` has the following arguments available for use:

```{r diagnostics-args, eval = FALSE, echo = TRUE}
args(diagnose)
```

If a directory is not provided (`dir = NULL`), then no output is saved at this point.

## Calculate an index of abundance

Finally, an index of abundance is estimated using `calc_index_areas()`

```{r index, eval = FALSE, echo = TRUE}
index <- calc_index_areas(
   data = fit$data,
   fit = fit,
   boundaries = "Coastwide"
)
```

`calc_index_areas()` has the following arguments available for use:


```{r index-args, eval = FALSE, echo = TRUE}
args(diagnose)
```

If a directory is not provided (`dir = NULL`), then no output is saved at this point.

## Write out all files and figures

Now that the model has been fit and diagnostics and an index of abundance has been calculated, all of the output objects can be written using `save_index_outputs()`:

```{r save-output, eval = FALSE, echo = TRUE}
save_index_outputs(
 fit = fit,
 diagnostics = diag,
 indices = index,
 dir = "C:/indices"
)
```

The directory structure created is:

```{r save-output-files, eval = FALSE, echo = TRUE}
 dir/
 └── species_name/
     └── survey_name/
         └── family_name/
             ├── data/
             │   ├── data.rdata
             │   └── fit.rds
             ├── diagnostics/
             │   ├── mesh.png
             │   ├── sanity_data_frame.csv
             │   ├── aic_nll.txt
             │   ├── run_diagnostics_and_estimates.rdata
             │   ├── qq.png
             │   ├── residuals_1_*.png
             │   ├── residuals_2_*.png (if delta model)
             │   ├── anisotropy.png
             │   ├── fixed_effects.png
             │   ├── density_*.png
             │   ├── data_with_residuals.rdata
             │   └── predictions.rdata
             └── index/
                 ├── est_by_area.csv
                 ├── index_coastwide.png (if Coastwide calculated)
                 └── index_all_areas.png
```

